---
title: "SST and DHW data processing"
author: "R. Cunning"
date: "2026-03-18"
output: html_document
---

```{r}
library(ncdf4)
library(raster)
library(terra)
library(tidyverse)
library(tidyterra)
library(sf)
library(rnaturalearth)
```

# Download Florida map
```{r}
# Download map for Florida
world <- rnaturalearth::ne_countries(scale = "large", returnclass = "sf")
# Create base map of Florida
basemap <- ggplot() +
  geom_sf(data = world, lwd = 0.1, fill = "gray90") +
  coord_sf(xlim = c(-83.2, -79.8), ylim = c(24.3, 27.3), expand = FALSE) +
  scale_fill_gradient2(high = "firebrick1", mid = "yellow", low = "forestgreen", 
                       midpoint = 0.5, limits = c(0, 1), na.value = NA,
                       labels = scales::label_percent(), name = "Corals\nbleached") +
  theme(text = element_text(size = 10),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "lightsteelblue1"),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid = element_blank(),
        legend.position = c(0.2, 0.5),
        legend.background = element_blank())


# Create reef hull/polygon surrounding all surveyed sites
all_survey_pts0 <- read_csv("data/drm/FRRP_EXPORT.csv") %>%
  distinct(Latitude, Longitude)

all_survey_pts <- sf::st_as_sf(x = all_survey_pts0, coords = c('Longitude', 'Latitude'))
reef_hull <- sf::st_concave_hull(sf::st_union(all_survey_pts), ratio = 0.09)
reef_hull <- sf::st_buffer(reef_hull, dist = 0.015)
reef_hull <- reef_hull %>% sf::st_set_crs(4326)
saveRDS(reef_hull, "data/spatial/reef_hull.rds")
all_survey_pts <- all_survey_pts %>% sf::st_set_crs(4326)



basemap +
  geom_sf(data = reef_hull, fill = NA, color = "black", linewidth = 0.7) +
  geom_sf(data = all_survey_pts, size = 1, alpha = 0.05) +
  coord_sf(xlim = c(-83.2, -79.8), ylim = c(24.3, 27.3), expand = FALSE)
```


# Download all SST data
```{r download_sst_data}

# Pixel-level CoralTemp SST stacks for Florida Reef Tract (robust lon/lat indexing + scaling)
library(tidyverse)
library(lubridate)
library(purrr)
library(ncdf4)
library(terra)
library(stringr)

# ---- CONFIG ----
base_sst <- "https://www.star.nesdis.noaa.gov/pub/socd/mecb/crw/data/5km/v3.1_op/nc/v1.0/daily/sst"
years <- c(1997:2025)
start_mmdd <- "03-01"
end_mmdd   <- "12-31"
out_dir <- "data/coraltemp"

# Florida Reef Tract bbox (WGS84)
bbox <- list(lon_min = -83.525, lon_max = -78.975, lat_min = 23.975, lat_max = 28.025)

dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)

# ---- BUILD DATE LIST (preserve Date class) ----
dates <- map(years, \(yy) seq.Date(as.Date(sprintf("%d-%s", yy, start_mmdd)),
                                   as.Date(sprintf("%d-%s", yy, end_mmdd)),
                                   by = "day")) |> 
  list_c()

# ---- URLS + DOWNLOAD (only missing files) ----
sst_urls  <- sprintf("%s/%d/coraltemp_v3.1_%s.nc", base_sst, year(dates), format(dates, "%Y%m%d"))
sst_files <- file.path(out_dir, basename(sst_urls))
miss <- which(!file.exists(sst_files))
#if (length(miss)) walk2(sst_urls[miss], sst_files[miss], \(u, f) download.file(u, f, mode = "wb"))

library(purrr)
library(furrr)
library(future)
# download in parallel
if (length(miss)) {
  plan(multisession, workers = 8) 
  future_walk2(
    sst_urls[miss], sst_files[miss],
    \(u, f) download.file(u, f, mode = "wb", quiet = TRUE)
  )
}
```

```{r}
# Download CRW DHW data
# DHW data downloaded 5/1/24 using:
# wget -r -l1 -A "ct*" https://www.star.nesdis.noaa.gov/pub/sod/mecb/crw/data/5km/v3.1_op/nc/v1.0/daily/dhw/2023/

```

```{r}
# ============================================================
# 1) Helpers + constants
# ============================================================

file_date <- function(f) {
  as.Date(stringr::str_match(basename(f), "(\\d{8})")[,2], "%Y%m%d")
}

# Florida Reef Tract extent
fl_ext <- terra::ext(-83.525, -78.975, 23.975, 28.025)

years_keep   <- c(2014, 2015, 2023, 2024)
start_mmdd   <- "06-01"
end_mmdd     <- "12-31"
cushion_days <- 83L   # to allow full 84-day windows at June 1

sst_dir   <- "data/coraltemp"
clim_file <- "data/coraltemp/ct5km_climatology_v3.1.nc"

# optional year-specific MMM offsets
add_by_year <- c("2014" = 0, "2015" = 0, "2023" = 0, "2024" = 0)
```

```{r}
# ============================================================
# 2) Read native MMM once and confirm native grid matches SST
# ============================================================

mmm_full <- rast(clim_file, subds = "sst_clim_mmm")

sst_files_all <- list.files(
  sst_dir,
  pattern = "^coraltemp_v3\\.1_\\d{8}\\.nc$",
  full.names = TRUE
)
sst_dates_all <- file_date(sst_files_all)

sst1 <- rast(sst_files_all[1], subds = "analysed_sst")

# confirm native geometry matches
stopifnot(compareGeom(sst1, mmm_full, stopOnError = FALSE))

# crop both to the same extent, preserving native grid
mmm0 <- crop(mmm_full, fl_ext)
sst1_crop <- crop(sst1, fl_ext)

# confirm cropped geometry still matches
stopifnot(compareGeom(sst1_crop, mmm0, stopOnError = FALSE))
```

```{r}
# ============================================================
# 3) Read CRW DHW for a year
# ============================================================

read_crw_dhw_year <- function(year, dhw_dir = "data/dhw", ext_use = fl_ext) {

  year_dir <- file.path(dhw_dir, as.character(year))
  files <- list.files(year_dir, pattern = "\\.nc$", full.names = TRUE)

  dates <- file_date(files)

  keep <- dates >= as.Date(sprintf("%d-%s", year, start_mmdd)) &
          dates <= as.Date(sprintf("%d-%s", year, end_mmdd))

  files <- files[keep]
  dates <- dates[keep]

  r <- rast(files, subds = "degree_heating_week")
  r <- crop(r, ext_use)

  # Match CRW style: noon UTC timestamps
  time(r) <- as.POSIXct(paste(dates, "12:00:00"), tz = "UTC")

  r
}
```

```{r}
# ============================================================
# 4) Compute DHW directly on native SST grid
# ============================================================

compute_dhw_rasters <- function(sst_stack, mmm_raster,
                                start_date, end_date,
                                mmm_add = 0,
                                set_time_noon_utc = TRUE) {

  # sanity
  stopifnot(compareGeom(sst_stack[[1]], mmm_raster, stopOnError = FALSE))

  # MMM adjustment if desired
  mmm_shift <- mmm_raster + mmm_add

  # HotSpot = SST - MMM
  hotspot <- sst_stack - mmm_shift

  # retain only HotSpot >= 1; otherwise 0
  hotspot_ge1 <- ifel(hotspot >= 1, hotspot, 0)

  # convert to daily DHW contribution in degree-C-weeks
  dhd <- hotspot_ge1 / 7

  # rolling 84-day sum, including current day
  n <- nlyr(dhd)
  w <- 84L
  out <- vector("list", n)

  for (i in seq_len(n)) {
    j0 <- max(1L, i - w + 1L)
    out[[i]] <- app(dhd[[j0:i]], sum, na.rm = TRUE)
  }

  dhw_stack <- rast(out)

  # preserve source dates from SST stack
  sst_dates <- as.Date(time(sst_stack))

  keep <- sst_dates >= start_date & sst_dates <= end_date
  dhw_stack <- dhw_stack[[which(keep)]]
  keep_dates <- sst_dates[keep]

  if (set_time_noon_utc) {
    time(dhw_stack) <- as.POSIXct(paste(keep_dates, "12:00:00"), tz = "UTC")
  } else {
    time(dhw_stack) <- keep_dates
  }

  names(dhw_stack) <- paste0("dhw_", keep_dates)

  peak_dhw <- app(dhw_stack, max, na.rm = TRUE)
  names(peak_dhw) <- "peak_dhw"

  list(
    dhw_stack = dhw_stack,
    peak_dhw  = peak_dhw
  )
}
```

```{r}
# ============================================================
# 5) Build calculated DHW stacks for all years
# ============================================================

dhw_stacks_calc <- list()
dhw_stacks_crw  <- list()
peak_calc       <- list()
peak_crw        <- list()

for (yy in years_keep) {

  message("Processing year: ", yy)

  start_date <- as.Date(sprintf("%d-%s", yy, start_mmdd))
  end_date   <- as.Date(sprintf("%d-%s", yy, end_mmdd))

  # include prior 83 days so June 1 has full 84-day accumulation window
  keep <- sst_dates_all >= (start_date - cushion_days) &
          sst_dates_all <= end_date

  sst_files <- sst_files_all[keep]
  sst_dates <- sst_dates_all[keep]

  if (length(sst_files) == 0) {
    stop("No SST files found for year ", yy)
  }

  # read SST directly on native CRW grid
  sst_stack <- rast(sst_files, subds = "analysed_sst")
  sst_stack <- crop(sst_stack, fl_ext)

  # preserve exact calendar date then write times like CRW
  time(sst_stack) <- as.POSIXct(paste(sst_dates, "12:00:00"), tz = "UTC")

  # confirm still aligned with MMM
  stopifnot(compareGeom(sst_stack[[1]], mmm0, stopOnError = FALSE))

  add_val <- add_by_year[[as.character(yy)]] %||% 0

  res <- compute_dhw_rasters(
    sst_stack   = sst_stack,
    mmm_raster  = mmm0,
    start_date  = start_date,
    end_date    = end_date,
    mmm_add     = add_val,
    set_time_noon_utc = TRUE
  )

  dhw_stacks_calc[[as.character(yy)]] <- res$dhw_stack
  peak_calc[[as.character(yy)]]       <- res$peak_dhw

  # also read CRW product
  dhw_stacks_crw[[as.character(yy)]] <- read_crw_dhw_year(yy, ext_use = fl_ext)
  peak_crw[[as.character(yy)]] <- app(dhw_stacks_crw[[as.character(yy)]], max, na.rm = TRUE)

  # convenience objects
  assign(paste0("dhw_stack_calc_", yy), res$dhw_stack, envir = .GlobalEnv)
  assign(paste0("peak_dhw_calc_", yy),  res$peak_dhw,  envir = .GlobalEnv)

  assign(paste0("dhw_stack_crw_", yy), dhw_stacks_crw[[as.character(yy)]], envir = .GlobalEnv)
  assign(paste0("peak_dhw_crw_", yy),  peak_crw[[as.character(yy)]],       envir = .GlobalEnv)
}
```

```{r}
# ============================================================
# 6) Comparison function
# ============================================================

compare_dhw_stacks <- function(calc_stack, crw_stack) {

  calc_dates <- as.Date(time(calc_stack))
  crw_dates  <- as.Date(time(crw_stack))

  stopifnot(identical(calc_dates, crw_dates))
  stopifnot(compareGeom(calc_stack[[1]], crw_stack[[1]], stopOnError = FALSE))

  diff_stack <- calc_stack - crw_stack

  stats <- global(diff_stack, c("min", "mean", "max"), na.rm = TRUE) |>
    as.data.frame() |>
    tibble::rownames_to_column("layer") |>
    mutate(date = calc_dates)

  frac_diff <- global(abs(diff_stack) > 1e-6, "mean", na.rm = TRUE)[,1]

  diff_ts <- tibble(
    date = calc_dates,
    mean_diff = global(diff_stack, "mean", na.rm = TRUE)[,1],
    min_diff  = global(diff_stack, "min",  na.rm = TRUE)[,1],
    max_diff  = global(diff_stack, "max",  na.rm = TRUE)[,1],
    frac_cells_different = frac_diff
  )

  peak_diff_max <- app(diff_stack, max, na.rm = TRUE)
  peak_diff_min <- app(diff_stack, min, na.rm = TRUE)

  list(
    diff_stack    = diff_stack,
    stats         = stats,
    diff_ts       = diff_ts,
    peak_diff_max = peak_diff_max,
    peak_diff_min = peak_diff_min
  )
}
```

```{r}
# ============================================================
# 7) Run comparisons
# ============================================================

cmp_2023 <- compare_dhw_stacks(dhw_stack_calc_2023, dhw_stack_crw_2023)
cmp_2024 <- compare_dhw_stacks(dhw_stack_calc_2024, dhw_stack_crw_2024)
cmp_2014 <- compare_dhw_stacks(dhw_stack_calc_2014, dhw_stack_crw_2014)
cmp_2015 <- compare_dhw_stacks(dhw_stack_calc_2015, dhw_stack_crw_2015)

cmp_2023$stats
cmp_2024$stats
cmp_2014$stats
cmp_2015$stats
```

Calculated DHW's match CRW's DHW product so we can move forward with the calculated values

# Plot peak calculated DHWs for each pixel for during each year
```{r}
# peak calculated dhw for each pixel in each year
peak_calc

#reef pixels
reef_hull_v <- terra::vect(reef_hull)

# Highest DHW reached in reef pixel in 2023?
maxdhw <- global(mask(peak_calc$`2023`, reef_hull_v), "max", na.rm = TRUE)[1,1]
# Lowest DHW reach in reef pixel in 2023
global(mask(peak_calc$`2023`, reef_hull_v), "min", na.rm = TRUE)[1,1]
# now want south of Miami 25.7617 lat
global(mask(mask(peak_calc$`2023`, reef_hull_v), init(peak_calc$`2023`, "y") < 25.7617, maskvalues = FALSE), "min", na.rm = TRUE)[1,1]

make_peak_map <- function(r, yy, reef_hull_v) {
  r_reef <- terra::mask(r, reef_hull_v)

  basemap +
    geom_spatraster(data = r_reef) +
    geom_sf(data = world, linewidth = 0.1, fill = "gray90") +
    coord_sf(xlim = c(-83.2, -79.8), ylim = c(24.3, 27.3), expand = FALSE) +
    scale_fill_distiller(
      palette = "Spectral",
      limits = c(0, maxdhw),
      na.value = NA,
      name = "Peak\nDHW "
    )
}

dhwmap_calc14 <- make_peak_map(peak_calc[["2014"]], "2014", reef_hull_v)
dhwmap_calc15 <- make_peak_map(peak_calc[["2015"]], "2015", reef_hull_v)
dhwmap_calc23 <- make_peak_map(peak_calc[["2023"]], "2023", reef_hull_v)
dhwmap_calc24 <- make_peak_map(peak_calc[["2024"]], "2024", reef_hull_v)

cowplot::plot_grid(
  dhwmap_calc14, dhwmap_calc15,
  dhwmap_calc23, dhwmap_calc24,
  ncol = 2,
  labels = c("2014", "2015", "2023", "2024"),
  label_x = 0.1, label_y = 0.92
)

ggsave(
  filename = "figures/FigS2.png",
  width = 200, height = 185, units = "mm"
)

```


# Extract DHWs for each surveyed site on date surveyed
```{r}

# ------------------------------------------------------------
# Helper: standardize raster CRS if terra reports "unknown"
# ------------------------------------------------------------
fix_lonlat_crs <- function(r) {
  r_crs <- terra::crs(r)
  if (is.na(r_crs) || r_crs == "" || grepl('GEOGCRS\\["unknown"', r_crs)) {
    terra::crs(r) <- "OGC:CRS84"
  }
  r
}

# great-circle distance in meters
haversine_m <- function(lon1, lat1, lon2, lat2) {
  rad <- pi / 180
  lon1 <- lon1 * rad
  lat1 <- lat1 * rad
  lon2 <- lon2 * rad
  lat2 <- lat2 * rad

  dlon <- lon2 - lon1
  dlat <- lat2 - lat1

  a <- sin(dlat / 2)^2 + cos(lat1) * cos(lat2) * sin(dlon / 2)^2
  2 * 6371000 * asin(pmin(1, sqrt(a)))
}

extract_layer_values_diag <- function(r_layer, df_xy, max_dist_m = 6000) {
  r_layer <- fix_lonlat_crs(r_layer)

  # first try: direct cell extraction
  simple <- terra::extract(
    r_layer,
    df_xy[, c("Longitude", "Latitude")],
    method = "simple",
    ID = FALSE
  )[, 1]

  vals <- terra::values(r_layer, mat = FALSE)
  good_cells <- which(!is.na(vals))

  good_xy <- terra::xyFromCell(r_layer, good_cells)
  good_vals <- vals[good_cells]

  final <- simple
  used_nearest <- rep(FALSE, length(simple))
  nearest_dist_m <- rep(NA_real_, length(simple))

  miss <- which(is.na(simple))

  for (i in miss) {
    d <- haversine_m(
      lon1 = df_xy$Longitude[i],
      lat1 = df_xy$Latitude[i],
      lon2 = good_xy[, 1],
      lat2 = good_xy[, 2]
    )

    j <- which.min(d)

    if (length(j) == 1 && is.finite(d[j])) {
      nearest_dist_m[i] <- d[j]

      if (d[j] <= max_dist_m) {
        final[i] <- as.numeric(good_vals[j])
        used_nearest[i] <- TRUE
      }
    }
  }

  tibble(
    value_simple = simple,
    value_final = final,
    used_nearest = used_nearest,
    nearest_dist_m = nearest_dist_m
  )
}

# ------------------------------------------------------------
# For a survey table with one row per survey, extract raster values
# from the layer matching each survey date
#
# Required columns in surveys:
#   Date, Longitude, Latitude
#
# value_name is the output column name, e.g. "dhw_crw"
# ------------------------------------------------------------
extract_stack_on_survey_dates <- function(rstk, surveys, prefix = "dhw", max_dist_m = 6000) {
  rstk <- fix_lonlat_crs(rstk)

  layer_dates  <- as.Date(time(rstk))
  survey_dates <- as.Date(surveys$Date)

  lyr_idx <- match(survey_dates, layer_dates)

  out_simple <- rep(NA_real_, nrow(surveys))
  out_final  <- rep(NA_real_, nrow(surveys))
  out_used   <- rep(FALSE, nrow(surveys))
  out_dist   <- rep(NA_real_, nrow(surveys))

  idx_by_date <- split(seq_len(nrow(surveys)), survey_dates)

  for (dd in names(idx_by_date)) {
    rows <- idx_by_date[[dd]]
    k <- match(as.Date(dd), layer_dates)

    if (is.na(k)) next

    ex <- extract_layer_values_diag(
      r_layer = rstk[[k]],
      df_xy = surveys[rows, c("Longitude", "Latitude")],
      max_dist_m = max_dist_m
    )

    out_simple[rows] <- ex$value_simple
    out_final[rows]  <- ex$value_final
    out_used[rows]   <- ex$used_nearest
    out_dist[rows]   <- ex$nearest_dist_m
  }

  surveys[[paste0(prefix, "_simple")]] <- out_simple
  surveys[[prefix]] <- out_final
  surveys[[paste0(prefix, "_used_nearest")]] <- out_used
  surveys[[paste0(prefix, "_nearest_dist_m")]] <- out_dist
  surveys$layer_matched <- lyr_idx

  surveys
}
```


```{r}
# ------------------------------------------------------------
# Read survey data once
# ------------------------------------------------------------
surveys_all <- read_csv("data/drm/FRRP_EXPORT.csv", show_col_types = FALSE) %>%
  mutate(
    Date = as_date(as.character(Date), format = "%m/%d/%Y %H:%M:%S")
  ) %>%
  filter(!is.na(Date),
         !is.na(Longitude),
         !is.na(Latitude)) %>%
  distinct(Site, Date, Latitude, Longitude, Subregion)
```


```{r}
# ------------------------------------------------------------
# Named lists of stacks
# ------------------------------------------------------------
dhw_stacks_crw <- list(
  "2014" = dhw_stack_crw_2014,
  "2015" = dhw_stack_crw_2015,
  "2023" = dhw_stack_crw_2023,
  "2024" = dhw_stack_crw_2024
)

dhw_stacks_calc <- list(
  "2014" = dhw_stack_calc_2014,
  "2015" = dhw_stack_calc_2015,
  "2023" = dhw_stack_calc_2023,
  "2024" = dhw_stack_calc_2024
)

years_keep <- c(2014, 2015, 2023, 2024)



survey_dhw_list <- vector("list", length(years_keep))
names(survey_dhw_list) <- as.character(years_keep)

for (yy in years_keep) {

  message("Extracting survey-date DHW for year: ", yy)

  surveys_yy <- surveys_all %>%
    filter(year(Date) == yy)

  if (nrow(surveys_yy) == 0) {
    warning("No surveys found for year ", yy)
    next
  }

  r_crw  <- dhw_stacks_crw[[as.character(yy)]]
  r_calc <- dhw_stacks_calc[[as.character(yy)]]

  # sanity checks
  stopifnot(identical(as.Date(time(r_crw)), as.Date(time(r_calc))))
  stopifnot(compareGeom(r_crw[[1]], r_calc[[1]], stopOnError = FALSE))

  # extract CRW
  out_yy <- extract_stack_on_survey_dates(
  rstk = r_crw,
  surveys = surveys_yy,
  prefix = "dhw_crw",
  max_dist_m = 6000
)

out_yy <- extract_stack_on_survey_dates(
  rstk = r_calc,
  surveys = out_yy,
  prefix = "dhw_calc",
  max_dist_m = 6000
) %>%
  mutate(
    year = yy,
    dhw_diff = dhw_calc - dhw_crw
  )

  survey_dhw_list[[as.character(yy)]] <- out_yy

  # write_csv(
  #   out_yy,
  #   file = sprintf("data/processed/%d_survey_date_dhw_compare.csv", yy)
  # )
}

survey_dhw_compare <- bind_rows(survey_dhw_list)

# write_csv(
#   survey_dhw_compare,
#   file = "data/processed/survey_date_dhw_compare_all_years.csv"
# )
```

## Compare extracted DHW values from CRW and from SST-calculations
```{r}
# overall agreement
survey_dhw_compare %>%
  summarize(
    n = n(),
    n_missing_crw = sum(is.na(dhw_crw)),
    n_missing_calc = sum(is.na(dhw_calc)),
    mean_diff = mean(dhw_diff, na.rm = TRUE),
    median_diff = median(dhw_diff, na.rm = TRUE),
    max_abs_diff = max(abs(dhw_diff), na.rm = TRUE),
    cor = cor(dhw_crw, dhw_calc, use = "complete.obs")
  )

# by year
survey_dhw_compare %>%
  group_by(year) %>%
  summarize(
    n = n(),
    mean_diff = mean(dhw_diff, na.rm = TRUE),
    median_diff = median(dhw_diff, na.rm = TRUE),
    max_abs_diff = max(abs(dhw_diff), na.rm = TRUE),
    cor = cor(dhw_crw, dhw_calc, use = "complete.obs"),
    .groups = "drop"
  )
```

# Write calculated DHWs for each survey site on date surveyed to file
```{r}

# format once
df_out <- survey_dhw_compare %>%
  dplyr::select(Date, Site, Subregion, Longitude, Latitude,
         dhw_crw, dhw_calc, year)

# loop years and write
for (yy in unique(df_out$year)) {

  df_yy <- df_out %>% filter(year == yy)

  # # CRW
  # df_crw <- df_yy %>%
  #   dplyr::select(Date, Site, Subregion, Longitude, Latitude, dhw = dhw_crw)
  # 
  # write_csv(
  #   df_crw,
  #   file = sprintf("data/processed/%d_dhw_crw_processed.csv", yy),
  #   na = ""
  # )

  # calculated
  df_calc <- df_yy %>%
    dplyr::select(Date, Site, Subregion, Longitude, Latitude, dhw = dhw_calc)

  write_csv(
    df_calc,
    file = sprintf("data/processed/%d_dhw_calc_site_surveyed.csv", yy),
    na = ""
  )
}
```

# Write full DHW trajectory for each surveyed site to file
```{r}
# ============================================================
# Write calculated DHW trajectories for every site on every day
# June 1–Dec 31, using only sites surveyed in that same year
# Output files like: 2014_dhw_calc_site_traj.csv
# ============================================================

extract_full_site_traj <- function(rstk, sites_df, max_dist_m = 6000) {
  rstk <- fix_lonlat_crs(rstk)

  layer_dates <- as.Date(time(rstk))
  out_list <- vector("list", length(layer_dates))

  for (k in seq_along(layer_dates)) {
    ex <- extract_layer_values_diag(
      r_layer = rstk[[k]],
      df_xy = sites_df[, c("Longitude", "Latitude")],
      max_dist_m = max_dist_m
    )

    out_list[[k]] <- sites_df %>%
      mutate(
        date = layer_dates[k],
        dhw  = ex$value_final
      ) %>%
      select(date, Site, Subregion, Longitude, Latitude, dhw)
  }

  bind_rows(out_list) %>%
    arrange(date)
}

for (yy in years_keep) {
  message("Writing full site DHW trajectory for year: ", yy)

  r_calc <- dhw_stacks_calc[[as.character(yy)]]

  if (is.null(r_calc)) {
    warning("No calculated DHW stack found for year ", yy)
    next
  }

  # only sites surveyed in this year
  sites_yy <- surveys_all %>%
    filter(lubridate::year(Date) == yy) %>%
    distinct(Site, Subregion, Longitude, Latitude)

  if (nrow(sites_yy) == 0) {
    warning("No survey sites found for year ", yy)
    next
  }

  df_yy <- extract_full_site_traj(
    rstk = r_calc,
    sites_df = sites_yy,
    max_dist_m = 6000
  )

  write_csv(
    df_yy,
    file = sprintf("data/processed/%d_dhw_calc_site_traj.csv", yy),
    na = ""
  )
}
```


